---
title: "HW3_WorkingFile_102020"
author: "JayGreene"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

library(tidyverse)
library(sf)
library(RSocrata)
library(viridis)
library(spatstat)
library(raster)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
# functions
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

```

# Read in Data from Chicago - Motor Vehicle Thefts

This uses the Socrata package for some data sets

```{r}
policeDistricts <- 
  st_read("https://data.cityofchicago.org/api/geospatial/fthy-xz3r?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = dist_num)
  
policeBeats <- 
  st_read("https://data.cityofchicago.org/api/geospatial/aerh-rz74?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = beat_num)

bothPoliceUnits <- rbind(mutate(policeDistricts, Legend = "Police Districts"), 
                         mutate(policeBeats, Legend = "Police Beats"))

mvtheft <-
  read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2017/d62x-nvdr") %>% 
    filter(Primary.Type == "MOTOR VEHICLE THEFT" & Description == "AUTOMOBILE") %>%
    mutate(x = gsub("[()]", "", Location)) %>%
    separate(x,into= c("Y","X"), sep=",") %>%
    mutate(X = as.numeric(X),Y = as.numeric(Y)) %>% 
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant")%>%
    st_transform('ESRI:102271') %>% 
    distinct()

chicagoBoundary <- 
  st_read(file.path(root.dir,"/Chapter5/chicagoBoundary.geojson")) %>%
  st_transform('ESRI:102271') 
```

## visualizing point data

Plotting point data and density- Points to Polygons

```{r fig.width=6, fig.height=4}
# uses grid.arrange to organize indpendent plots
grid.arrange(ncol=2,
ggplot() + 
  geom_sf(data = chicagoBoundary) +
  geom_sf(data = mvtheft, colour="red", size=0.1, show.legend = "point") +
  labs(title= "Motor Vehicle Theft, Chicago - 2017") +
  mapTheme(title_size = 14),

ggplot() + 
  geom_sf(data = chicagoBoundary, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(mvtheft)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_viridis() +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Motor Vehicle Theft") +
  mapTheme(title_size = 14) + theme(legend.position = "none"))
```

## Creating a fishnet grid

A Fishnet is a grid where every polygon is connected on all sides by polygon neighbors continuously. We can define for every polygon what its neighbors are. Allows us to run analytics.


The `{sf}` package offers really easy way to create fishnet grids.

```{r}
## using {sf} to create the grid
## Note the `.[chicagoBoundary] %>% ` line. This is needed to clip the grid to our data
fishnet <- 
  st_make_grid(chicagoBoundary,
               cellsize = 500, 
               square = TRUE) %>%
  .[chicagoBoundary] %>% 
  st_sf() %>%
  mutate(uniqueID = rownames(.))
```

### Aggregate points to the fishnet

> How can we aggregate points into a fishnet grid?

```{r}
## add a value of 1 to each crime, sum them with aggregate
crime_net <- 
  dplyr::select(mvtheft) %>% 
  mutate(countmvtheft = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countmvtheft = replace_na(countmvtheft, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

ggplot() +
  geom_sf(data = crime_net, aes(fill = countmvtheft), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of MVTheft for the fishnet") +
  mapTheme()

# For demo. requires updated mapview package
# xx <- mapview::mapview(crime_net, zcol = "countBurglaries")
# yy <- mapview::mapview(mutate(burglaries, ID = seq(1:n())))
# xx + yy
```
## Modeling Spatial Features

```{r}
## only pulling a single variable for our model to keep it simple
## using Socrata again
abandonCars <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Abandoned-Vehicles/3c9v-pnva") %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2017") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Cars")

abandonBuildings <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Vacant-and-Abandoned-Building/7nii-7srd") %>%
    mutate(year = substr(date_service_request_was_received,1,4)) %>%  filter(year == "2017") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Buildings")

graffiti <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Graffiti-Removal-Historical/hec5-y4x5") %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2017") %>%
    filter(where_is_the_graffiti_located_ %in% c("Front", "Rear", "Side")) %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Graffiti")

streetLightsOut <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Street-Lights-All-Out/zuxi-7xem") %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2017") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Street_Lights_Out")

sanitation <-
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Sanitation-Code-Complaints-Hi/me59-5fac") %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2017") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Sanitation")

liquorRetail <- 
  read.socrata("https://data.cityofchicago.org/resource/nrmj-3kcf.json") %>%  
    filter(business_activity == "Retail Sales of Packaged Liquor") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Liquor_Retail")

traffic <-
  read.socrata("https://data.cityofchicago.org/resource/pfsx-4n4m.json")





## Neighborhoods to use in LOOCV in a bit
neighborhoods <- 
  st_read("https://raw.githubusercontent.com/blackmad/neighborhoods/master/chicago.geojson") %>%
  st_transform(st_crs(fishnet)) 

```

### Aggregating Features to Fishnet

```{r}

vars_net <- abandonCars %>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
  full_join(fishnet, by = "uniqueID") %>%
  spread(Legend, count, fill=0) %>%
  st_sf() %>%
  dplyr::select(-`<NA>`) %>%
  na.omit() %>%
  ungroup()

# vars_net <- abandonCars %>%
#   spatially join abandonCars points to the fishnet polygon they are within %>%
#   drop the geometry attribute %>%
#   group_by each cells ID and the name of the feature %>%
#   summarize count the number of each point per grid cell %>%
#   join that summary back to spatial fishnet by cell ID %>%
#   "spread" from long to wide format and make column of our point count %>%
#   tell R that this should be an sf object %>%
#   remove a fussy column that appears b/c of NA %>%
#   get rid of rows with an NA in any column %>%
#   remove grouping so you are not tripped up later
```
